Choosing climate forcing for numerical simulations of the Cordilleran ice sheet

### 2013-02-17 sections/model.tex
Added subsection 3.2 Ice dynamics

%\subsection{Ice flow dynamics}
%To resolve ice dynamics, PISM specifically uses a heuristic combination of two classical and well-based zeroth-order shallow-ice approximations through a weighting function \citep{bueler-brown-2009}. The Shallow Shelf Approximation \citep[SSA,][]{morland-1987,weis-etal-1999} is used as a ``sliding law'' for the Shallow Ice Approximation \citep[SSA,][]{hutter-1983}. Depth-averaged velocities $\vec{v}$ are computed as a weighted sum of SIA velocities $\vec{v_{\mathrm{SIA}}}$ and SSA velocities $\vec{v_{\mathrm{SSA}}}$ by
%
%\begin{equation}
%	\vec{v} = f(|\vec{v_{\mathrm{SSA}}}|)\vec{v_{\mathrm{SIA}}}
%	+ (1-f(|\vec{v_{\mathrm{SSA}}}|))\vec{v_{\mathrm{SSA}}},
%\end{equation}
%
%with
%
%\begin{equation}
%	f(|\vec{v_{\mathrm{SSA}}}|) = 1
%	- \frac{2}{\pi}\arctan{\frac{|\vec{v_{\mathrm{SSA}}}|^2}{100^2}}.
%\end{equation}
%
%This approach, although using a heuristic, allows realistic representation of horizontal shear and longitudinal stresses where they matter \citep{bueler-brown-2009}.
%
%
%The SSA velocities are determined by a a pseudo-plastic sliding law in which the yield stress is depends on the availability of basal water and a prescibed till friction angle. We apply lower friction angles at lower elevation to model more slippery beds due to the presence of marine sediments.
%
%\begin{equation}
%	\phi = \left\{\begin{array}{llc}
%		10\,\unit{\degree C} & \mathrm{for} &               z<  0\,\unit{m} \\
%		z/10 + 10°           & \mathrm{for} &   0\,\unit{m}<z<200\,\unit{m} \\
%		30\,\unit{\degree C} & \mathrm{for} & 200\,\unit{m}<z               \\
%	\end{array}\right.
%\end{equation}
%
%\subsection{Glacier bed interface}
%
%The ice-sheet bed topography responds to the ice-loads following a regional isostasy model with flexure and relaxation.
%
%Geothermal heat flux is prescribed to 70\,\unit{W\,m^{-2}}
%
%\subsection{Simulation design and numerical implementation}
%
%The simulations were run with PISM, a parallel ice sheet model (version stable 0.5.11). Documentation on the code can be found in several publications \citep[e.g.,][]{bueler-brown-2009,aschwanden-etal-2012} and more extensively online \citep[{\url{http://www.pism-docs.org}},][]{web:pism}.

>>>>

\subsection{Ice dynamics}

Ice flow velocities are computed by a heuristic where the Shallow Shelf Approximation (SSA) is used as a ``sliding law'' to a polythermal Shallow Ice Approximation (SIA) \citep{bueler-brown-2009}.

\todo[inline]{
	\textbf{Include SSA+SIA \citet{bueler-brown-2009} formula}

	Depth-averaged velocities $\vec{v}$ are computed as a weighted sum of SIA velocities $\vec{v_{\mathrm{SIA}}}$ and SSA velocities $\vec{v_{\mathrm{SSA}}}$ by

	\begin{equation}
		\vec{v} = f(|\vec{v_{\mathrm{SSA}}}|)\vec{v_{\mathrm{SIA}}}
	  + (1-f(|\vec{v_{\mathrm{SSA}}}|))\vec{v_{\mathrm{SSA}}},
	\end{equation}

	with

	\begin{equation}
		f(|\vec{v_{\mathrm{SSA}}}|) = 1
		- \frac{2}{\pi}\arctan{\frac{|\vec{v_{\mathrm{SSA}}}|^2}{100^2}}.
	\end{equation}
}

SSA velocities are determined by a a pseudo-plastic sliding law where the yield stress depends on the availability of basal water and a prescibed till friction angle$\phi$. We apply lower friction angles at low elevation to mimick the more slippery beds associated to the presence of marine sediments:

\begin{equation}
	\phi = \left\{\begin{array}{llc}
		10\,\unit{\degree C} & \mathrm{for} &               z<  0\,\unit{m} \\
		z/10 + 10°           & \mathrm{for} &   0\,\unit{m}<z<200\,\unit{m} \\
		30\,\unit{\degree C} & \mathrm{for} & 200\,\unit{m}<z               \\
	\end{array}\right.
\end{equation}

%\subsection{Glacier bed interface}
%
%The ice-sheet bed topography responds to the ice-loads following a regional isostasy model with flexure and relaxation.
%
%Geothermal heat flux is prescribed to 70\,\unit{W\,m^{-2}}
%
%\subsection{Simulation design and numerical implementation}
%
%The simulations were run with PISM, a parallel ice sheet model (version stable 0.5.11). Documentation on the code can be found in several publications \citep[e.g.,][]{bueler-brown-2009,aschwanden-etal-2012} and more extensively online \citep[{\url{http://www.pism-docs.org}},][]{web:pism}.
